### 2021-03-05 ch/2-related-work.tex
Incorporate Novi's feedback for chapter 1
Use British spelling

\chapter{Related work}\label{ch:related-work}
%
% =========================
%      old introduction
% =========================
% \section{Introduction}\label{introduction}
% Fairness in Machine Learning is a topic that has recently received increasing attention.
% Machine Learning has become very widespread in our society
% which brings some dangers with it.
% Decisions such as credit eligibility and hiring are increasingly done automatically.
% However, the algorithms that are used in these decisions can have strong biases
% that stem from the training data.
% Often, the algorithms are even more biased than the training data.
% \citet{kamishima2012fairness} give a great illustration for this
% (adapted from \citet{calders2010three}).
% The illustration is based on the Adult/Census Income dataset \citep{kohavi1996scaling}
% which contains records of individuals with thirteen attributes, among them age and gender.
% The task is to predict whether or not the individual earns more than \$50K per year.
% In the dataset, 32\% of the men have a positive label (\ie, they earn more than \$50K),
% but of the women only 11\% have a positive label.
% With this ratio of about 3:1, the dataset is already skewed.
% However, when training a \href{https://en.wikipedia.org/wiki/Naive_Bayes_classifier}{Na\"ive Bayes classifier} on the data,
% the ratio in the predictions is even more unfair:
% the amount of women who get a positive prediction is even lower than 11\%.
% If the Na\"ive Bayes classifier does not get access to the gender information,
% the predictions will be a bit fairer,
% but they are still more unfair than the original dataset labels.
%
% This phenomenon of predictions being more unfair than an already skewed dataset
% is due to the generalization that classifiers have to make.
% It is easier to classify all but the most outstanding women as low income
% than to learn a complicated decision boundary.
% Furthermore, removing the gender information does not solve the problem
% because the gender can be inferred from other data.
%
% The following is a review of the literature on Fairness in Machine Learning.
% Several different approaches to and definition of Fairness are presented.
% The focus is on identifying the major trends in the field from the beginning.
\section{\NoCaseChange{Two views of the dataset bias problem}}\label{two-views-of-the-dataset-bias-problem}
As hinted to in the introduction,
there is a divide in the relevant literature in how the problem of dataset bias is approached.
On the one hand, there is the ``ground-truth-centric'' view -- expressed through most of the introduction --
that the training set is biased but the deployment setting is not.
By \emph{deployment setting} I am here referring to the real distribution
that will be encountered by the machine learning system once it is in use.

@@

After the general setup, the authors define a concrete measure of discrimination:
\begin{align}
  \label{eq:disc}
  Disc := P(y=1|s =1) - P(y=1|s = 0)
\end{align}
(The paper uses a different notation.
The notation here reflects the notation in later papers.)

@@

Another approach that aims to define a better proxy for the fairness criteria is \citet{quadrianto2017recycling}.
In that paper, \emph{distribution matching} is used to evaluate fairness.
More specifically, like \citet{louizos2016variational}, Maximum Mean Discrepancy (MMD) is used to check
if the distributions \(P(\hat{y}|s=0)\) and \(P(\hat{y}|s=1)\) are close, to ensure Demographic Parity.
This can also be done for Equality of Opportunity and Equalized Odds with the appropriate distributions.
\citet{quadrianto2017recycling} show that the approach by \citet{zafar2017fairnessconstraints} is a special case of this.
Furthermore, in contrast to that approach,
the sensitive attribute is explicitly made available to the classifier during training,
but not during decision time (after training is finished).
This is done via \emph{privileged learning}.
Privileged learning is a variant of \acp{SVM}
where certain features are only available during training but not during decision time.
This is usually used in the context where some features are costly or impractical to gather.
An example is using (only during training) thermal images in addition to normal images for face recognition.
The additional information is intended to enable the classifier to learn the desired task quicker or better.

Combining privileged learning and the MMD fairness criterion,
the task becomes a multi-objective optimization for an SVM.
The MMD criterion is generally non-convex but there are several ways to deal with this.
The authors find the best algorithm (low discrimination, high accuracy)
with an evolutionary multi-objective optimization.
The evolutionary algorithm produces a number of variants of the classifier
with different trade-offs between classification error and discrimination.
As in \citet{zafar2017fairnesstreatment} before,
the Adult/Census Income dataset is used to test Demographic Parity
and the ProPublica/COMPAS dataset is used for Equality of Opportunity and Equalized Odds.
\citet{zafar2017fairnesstreatment} was used as a baseline
but the implementation failed on the Adult/Census Income dataset.
On the ProPublica/COMPAS dataset,
the proposed algorithm performed within error just as well as the fair baseline algorithm.

@@

\citet{kim2019learning} also consider the problem of learning from biased data (and evaluating on unbiased data).
While they do not formulate the problem as a fairness problem,
it is equivalent to enforcing fairness in the presence of severe sampling bias.
Their main motivating example is the Coloured MNIST dataset:
in this dataset, digits are randomly coloured in the test set,
but in the training set there is a one-to-one correspondence between digit class and colour.
A na\"ive classifier will learn to predict colour instead of digit class.
The proposed method is very similar to \citet{ganin2016domain} and \citet{edwards2016censoring},
if we think of the colours as forming different domains
(\ie, the \emph{red} domain contains all red digits, etc).
The goal is then to learn a domain-independent representation.
The main difference to previous work is an additional entropy loss term in the adversarial loss.

>>>>

%
% =========================
%      old introduction
% =========================
% \section{Introduction}\label{introduction}
% Fairness in Machine Learning is a topic that has recently received increasing attention.
% Machine Learning has become very widespread in our society
% which brings some dangers with it.
% Decisions such as credit eligibility and hiring are increasingly done automatically.
% However, the algorithms that are used in these decisions can have strong biases
% that stem from the training data.
% Often, the algorithms are even more biased than the training data.
% \citet{kamishima2012fairness} give a great illustration for this
% (adapted from \citet{calders2010three}).
% The illustration is based on the Adult/Census Income dataset \citep{kohavi1996scaling}
% which contains records of individuals with thirteen attributes, among them age and gender.
% The task is to predict whether or not the individual earns more than \$50K per year.
% In the dataset, 32\% of the men have a positive label (\ie, they earn more than \$50K),
% but of the women only 11\% have a positive label.
% With this ratio of about 3:1, the dataset is already skewed.
% However, when training a \href{https://en.wikipedia.org/wiki/Naive_Bayes_classifier}{Na\"ive Bayes classifier} on the data,
% the ratio in the predictions is even more unfair:
% the amount of women who get a positive prediction is even lower than 11\%.
% If the Na\"ive Bayes classifier does not get access to the gender information,
% the predictions will be a bit fairer,
% but they are still more unfair than the original dataset labels.
%
% This phenomenon of predictions being more unfair than an already skewed dataset
% is due to the generalization that classifiers have to make.
% It is easier to classify all but the most outstanding women as low income
% than to learn a complicated decision boundary.
% Furthermore, removing the gender information does not solve the problem
% because the gender can be inferred from other data.
%
% The following is a review of the literature on Fairness in Machine Learning.
% Several different approaches to and definition of Fairness are presented.
% The focus is on identifying the major trends in the field from the beginning.
\section{\NoCaseChange{Two views of the dataset bias problem}}\label{two-views-of-the-dataset-bias-problem}
As hinted to in the introduction,
there is a divide in the relevant literature in how the problem of dataset bias is approached.
On the one hand, there is the ``ground-truth-centric'' view -- expressed through most of the introduction --
that the training set is biased but the deployment setting is not.
By \emph{deployment setting} I am here referring to the real distribution
that will be encountered by the machine learning system once it is in use.

@@

The authors then define a concrete measure of discrimination:
\begin{align}
  \label{eq:disc}
  Disc := P(y=1|s =1) - P(y=1|s = 0)
\end{align}
% (The paper uses a different notation.
% The notation here reflects the notation in later papers.)

@@

Another approach that aims to define a better proxy for the fairness criteria is \citet{quadrianto2017recycling}.
In that paper, \emph{distribution matching} is used to evaluate fairness.
More specifically, like \citet{louizos2016variational}, \acf{MMD} is used to check
if the distributions \(P(\hat{y}|s=0)\) and \(P(\hat{y}|s=1)\) are close, to ensure Demographic Parity.
This can also be done for Equality of Opportunity and Equalized Odds with the appropriate distributions.
\citet{quadrianto2017recycling} show that the approach by \citet{zafar2017fairnessconstraints} is a special case of this.
Furthermore, in contrast to that approach,
the sensitive attribute is explicitly made available to the classifier during training,
but not during decision time (after training is finished).
This is done via \emph{privileged learning}.
Privileged learning is a variant of \acp{SVM}
where certain features are only available during training but not during decision time.
This is usually used in the context where some features are costly or impractical to gather.
An example is using (only during training) thermal images in addition to normal images for face recognition.
The additional information is intended to enable the classifier to learn the desired task quicker or better.

Combining privileged learning and the \ac{MMD} fairness criterion,
the task becomes a multi-objective optimization for an SVM.
The MMD criterion is generally non-convex but there are several ways to deal with this.
The authors find the best algorithm (low discrimination, high accuracy)
with an evolutionary multi-objective optimization.
The evolutionary algorithm produces a number of variants of the classifier
with different trade-offs between classification error and discrimination.
As in \citet{zafar2017fairnesstreatment} before,
the Adult/Census Income dataset is used to test Demographic Parity
and the ProPublica/COMPAS dataset is used for Equality of Opportunity and Equalized Odds.
\citet{zafar2017fairnesstreatment} was used as a baseline
but the implementation failed on the Adult/Census Income dataset.
On the ProPublica/COMPAS dataset,
the proposed algorithm performed within error just as well as the fair baseline algorithm.

@@

\citet{kim2019learning} also consider the problem of learning from biased data (and evaluating on unbiased data).
While they do not formulate the problem as a fairness problem,
it is equivalent to enforcing fairness in the presence of severe sampling bias.
Their main motivating example is the Coloured MNIST dataset:
in this dataset, digits are randomly coloured in the test set,
but in the training set there is a one-to-one correspondence between digit class and colour.
A na\"ive classifier will learn to predict colour instead of digit class.
The proposed method is very similar to \citet{ganin2016domain} and \citet{edwards2016censoring},
if we think of the colours as forming different domains
(\ie, the \emph{red} domain contains all red digits, etc.).
The goal is then to learn a domain-independent representation.
The main difference to previous work is an additional entropy loss term in the adversarial loss.
